Nonlinear mixed effects models for time series forecasting of smart meter demand

### 2020-09-29 paper.Rmd
Resolve #5
- Clarify that only 15 minute electricity demand data are being used from smart meters.
- Explain motivation behind using SSCs better. Clarify that curve refers to nonlinear demand response function.

This paper explores how electricity forecasting accuracy can be improved by using mixed effects models. We examine if mixed models can produce forecasts as accurately as separate models fit for each subject. We approach the problem in the context of producing one-day ahead forecasts of electricity demand for `r n_buildings` commercial office buildings in Australia. When working with mixed effects models, each building is treated as a random effect and building characteristics are treated as fixed effects. We attempt to model the relationship between temperature and demand using both linear and spline based methods.

@@

Several papers have shown the relationship between electricity demand and temperature are well modelled using nonparametric components such as cubic splines [@Hyndman2010-ui; @Fan2012-bs]. This paper uses a similar approach within a mixed model framework. Other papers that explore semiparametric mixed models include @Grajeda2016-vr; @Ugarte2009-hx; and @Durban2005-lk. @Durban2005-lk is of particular note as it introduces the concept of subject-specific curves using piecewise linear splines for longitudinal data. We build on the idea of subject-specific curves by applying them to time series data and incorporating natural splines.

@@

Smart meter data recorded at 15-minute intervals for `r n_buildings` buildings are used when training and validating our models. The electricity demand is normalised by each building's net lettable area (NLA) to ensure demand is comparable between buildings. An example of a day of smart meter readings from six buildings is shown in Figure \@ref(fig:wh-line). Temperature data recorded at 15-minute intervals from the closest available weather station are also available for each building.

The relationship between current temperature and electricity demand is shown in Figure \@ref(fig:temp-wh-scatter) for two buildings at midday and midnight. There is a clear difference in this relationship between the business and non-business periods. Furthermore, this relationship varies between each building. Note that when splines are fitted independently for each building they can be very sensitive to outliers or data at domain extremes. This partly motivates our use of mixed effects models where data from multiple buildings can be used to fit these relationships resulting in models that are more robust to outliers or sparse data at extremes.

>>>>

This paper explores how electricity forecasting accuracy can be improved by using mixed effects models. We examine if mixed models can produce forecasts as accurately as separate models fit for each subject. We approach the problem in the context of producing one-day ahead forecasts of electricity demand for `r n_buildings` commercial office buildings in Australia. Electricity demand readings at 15-minute intervals are taken from smart meters located at each of the buildings. These readings exhibit within-subject correlation which are well suited to modelling using mixed effects models, which we explore here. When working with mixed effects models, each building is treated as a random effect and building characteristics are treated as fixed effects. We attempt to model the relationship between temperature and demand using both linear and spline based methods.

@@

Several papers have shown the relationship between electricity demand and temperature are well modelled using nonparametric components such as cubic splines [@Hyndman2010-ui; @Fan2012-bs]. This paper uses a similar approach within a mixed model framework. Other papers that explore semiparametric mixed models include @Grajeda2016-vr; @Ugarte2009-hx; and @Durban2005-lk. @Durban2005-lk is of particular note as it introduces the concept of subject-specific curves using piecewise linear splines for longitudinal data. These subject specific curves are nonlinear functions that are allowed to vary by each subject. In this paper we first justify the use of subject-specific curves for modelling metered electricity demand before building on the existing methodology using natural splines.

@@

Electricity demand data from `r n_buildings` commercial office buildings in Australia are used to train and validate our models. The electricity demand data is recorded at 15-minute intervals by smart meters located at each building. The electricity demand is normalised by each building's net lettable area (NLA) to ensure demand is comparable between buildings. An example of a day of smart meter readings from six buildings is shown in Figure \@ref(fig:wh-line). Temperature data recorded at 15-minute intervals from the closest available weather station are also available for each building.

The relationship between current temperature and electricity demand is shown in Figure \@ref(fig:temp-wh-scatter) for two buildings at midday and midnight. There is a clear difference in this relationship between the business and non-business periods. Furthermore, this relationship varies between each building. The different relationships for each building necessitates the use of a nonlinear demand response function that can adequately capture the unique response characteristics of each building. To do this, we model our data with mixed effects models and capture the nonlinear building specific demand response using subject-specific curves. Note that when splines are fitted independently for each building they can be very sensitive to outliers or data at domain extremes. This also motivates our use of mixed effects models where data from multiple buildings can be used to fit these relationships giving models that are more robust to outliers or sparse data at covariate extremes.
